Making Soup: Preparing and Validating Molecular Simulations of the Bacterial Cytoplasm

### 2019-09-25 paper/virtual_cytoplasm_v2.tex
I'm working on the introduction

Biomolecules move and function in an environment densely packed with high concentrations of macromolecules. The presence of macromolecules leads to steric effect due to excluded volume effect and intermolecular attrative/repulsive forces due to distributed charges on the surface of macromolecules.

Structure and dynamics of biomolecules are well characterized {\em in vitro}, however these studies {\em in vivo} are still evolving. 
Investigating biomolecular properties {\em in vivo} is possible through developments in the fields of nuclear magnetic resonance~\cite{reckel2007,pielak2008}, or fluoroscence spectroscopies~\cite{ignatova2004,xie2008,English2011}.

An alternative method is to use computational models and simulation techniques. Biomolecular simulations are often carried out under dilute conditions or simple models of macromolecular crowdings~\cite{Spiga2014a,Harada2012a,Nawrocki2017a}. However, more attemps in modeling bacterial cytoplasm have been made recently~\cite{Mcguffee2010,Yu2016a}.

\begin{itemize}
   \item How a computational model can be useful
   \item From a simple extrapolation in time, it has been estimated that the simulation of an {\em E. coli} bacterium would be possible in 25 years from now for one nanosecond with $10^{11}$ atoms~\cite{vanGunsteren2006a}
\end{itemize}

Here we report on a model of {\em Escherichia coli} cytoplasm at atomistic level. The challenges in modeling and simulation with Molecular Dynamics are pointed and discussed and solutions are provided.

This work is the first model of {\em E. coli} at atomistic resolution that spans cellular dynamics on a microsecond scale.

@@

\begin{figure}[H]
\includegraphics[scale=0.5]{msd.pdf}
\includegraphics[scale=0.5]{diff_cyto_over_singles.pdf}
\caption{A)Diffusion coefficient of crowders from single molecule simulations (blue bars) and cytoplasm simulations (orange bars). B) The ratio between diffusion coefficient obtained from cytoplasm to single molecule simulations. The crowders are sorted according to their sizes on the x-axis.}
\label{fig:translational_diffusion}
\end{figure}

@@

\begin{figure}[H]
\includegraphics[scale=0.5]{rdf_RNA_metabolites.pdf}
\caption{Radial distribution function showing the probability of finding metabolites or cations around RNA molecules in the cytoplasm model.\color{red}\textbf{I SUBMITTED NEW RDF CALCULATIONS CENTERED AROUND tRNA BECAUSE I DIDN'T FIND THE XVG FILES THAT WERE USED TO GENERATE THIS PLOT}}
\label{fig:tRNA_aggregation}
\end{figure}

@@

\caption{A) Solvent accessible surface area (sasa) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars). The average sasa for each protein in single molecule simulation is over number of chains of proteins and three replica. The average sasa for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica. The error bars show the standard errors. B) Root mean squared displacement (rmsd) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars). The average rmsd for each protein in single molecule simulation is over number of chains of proteins and three replica. The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica. The error bars show the standard errors.}
\label{fig:structural_integrity_chain}
\end{figure}

@@

\begin{figure}[H]
\includegraphics[scale=0.5]{rmsf.pdf}
\caption{Root mean squared fluctuations (rmsf) for 8 crowders from single molecule simulation (blue bars) and cytoplasm simulation (orange bars). x-axis shows the residues for each chain of proteins. The average rmsf for each protein in single molecule simulation is over number of chains of proteins and three replica. The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica. The error bars show the standard errors.\textbf{\color{red}{CAN YOU ADD THE PROTEIN NAME AS A LABEL INSIDE EACH PLOT? I THINK THE VARIABLE NAME SHOULD BE IN THE Y-AXIS (RMSF NM2)}}}
\label{fig:rmsf}
\end{figure}

@@

It has been demonstrated that each of the force fields that are most widely used to perform molecular dynamics simulations of macromolecules have different results for highly concentrated amino acid solutions ~ \cite{andrews2013molecular}. Previous all-atom simulations of high scale cytoplasm models used parameters from the CHARMM family of force fields ~\cite{yu2016biomolecular,mcguffee2010diffusion}. In this work we use, for the first time, the AMBER force fields to study crowding effects on a cytoplasm model. All-atom explicit solvent molecular dynamics simulations have been reported for large-scale cytoplasm models of \textit{Mycoplasma genitalium} composed by 103 and 12 million atoms that were simulated for 20 and 60 \SI{}{\nano\second}, respectively ~\cite{yu2016biomolecular}. The cytoplasm model described in this work has 1.5 million atoms and was simulated for the total of \SI{3}{\micro\second}, an unprecedented time scale to study crowding effects on a heterogeneous cytoplasm model in atomic detail.

>>>>

%Biomolecules move and function in an environment densely packed with high concentrations of macromolecules. The presence of macromolecules leads to steric effect due to excluded volume effect and intermolecular attrative/repulsive forces due to distributed charges on the surface of macromolecules.

Crowding ... ~\cite{feig2017crowding} ~\citep{ostrowska2019modeling}

-> Crowding affects hydration around proteins ~\cite{Harada2012a,king2013crowding}
-> leads to stabilization of native structure due to the excluded volume effect ~\cite{cheung2005molecular}
-> can also destabilize proteins due to  ~\cite{feig2011variable,miklos2011protein,wang2012macromolecular}

While the structure and dynamics of biomolecules is well characterized \textit{in vitro}, the understanding of the influences of crowded environments for the \textit{in vivo} are still evolving.  Techniques such as nuclear magnetic resonance~\cite{reckel2007,pielak2008} and fluoroscence spectroscopy~\cite{ignatova2004,xie2008,English2011} are among the most relevant to the field due to their ability to probe cells in a non destructive way. Another essential class of methods come from computational modeling and simulation. By extrapolating the advances in computational power in the last decades to the future, it has been estimated that the atomistic simulation of an entire \textit{Escherichia coli} bacterium for \SI{1}{\nano\second} will be possible in 15 years from now ~\cite{vanGunsteren2006a}.

As the available computational power increases, models become increasingly complex and are better representations of reality. The most simple class of computational models apply brownian dynamics calculations to probe macromolecules imersed in a solution of crowders that are treated as hard spheres with varying radius to represent different types of molecules ~\cite{ando2010crowding}. These models usually use an implicit representation of the solvent ~\cite{mcguffee2010diffusion}.More recently, all-atom explicit solvent molecular dynamics simulations have been employed on crowded systems composed by several copies of one or a few macromolecules ~\cite{wang2017influence}. However, these models doesn't reflect the true heterogeneity of biologically relevant crowded environments such as the cytoplasm. This was addressed by the development of a detailed model for the cytoplasm of the bacterium \textit{Mycoplasma genitalium} containing 103 million atoms ~\cite{feig2015complete}. This model, as well as a smaller version with 13 million atoms, was submitted to molecular dynamics simulations ~\cite{yu2016biomolecular}. Despite the size of these systems, the simulation times were too short, \SI{20}{\nano\second} for the complete model and \SI{140}{\nano\second} for the smaller version, to accurately characterize the effects of the slow diffusion of its constituents.

In this work we report an all-atom model of the cytoplasm of \textit{Escherichia coli} that was built to reflect the real biological system. We discuss the difficulties that arise when building such system and provide a set of Python scripts and files that anyone can use to build its own crowded systems with custom parameters and elements. Additionally, we performed long molecular dynamics simulations of our cytoplasm model, reaching \SI{3}{\micro\second} of total simulation time. This work is the first to explore the structural dynamics of an all-atom explicit solvent cytoplasm model with microsecond scale molecular dynamics simulations.

@@

\begin{figure}[H]
\includegraphics[scale=0.5]{msd.pdf}
\includegraphics[scale=0.5]{diff_cyto_over_singles.pdf}
\caption{A)Diffusion coefficient of crowders from single molecule simulations (blue bars) and cytoplasm simulations (orange bars). B) The ratio between diffusion coefficient obtained from cytoplasm to single molecule simulations. The crowders are sorted according to their sizes on the x-axis. \textbf{\color{red}{THE Dcyto/Ddil PLOT IS "SQUEEZED" BECAUSE OF ITS STRANGE SCALE. IT WOULD BE BETTER TO CENTER THE DATAPOITS, E.G. USE THE Y-AXIS RANGE OF 0.0-0.2 OR SOMETHING SIMILAR. THIS SCALE WILL MAKE THE DIFFERENCE OF THE BEHAVIOR OF tRNA MORE VISIBLE AS WELL}}}
\label{fig:translational_diffusion}
\end{figure}

@@

\begin{figure}[H]
\includegraphics[scale=0.5]{rdf_RNA_metabolites.pdf}
\caption{Radial distribution function showing the probability of finding metabolites or cations around RNA molecules in the cytoplasm model.\color{red}\textbf{THIS PLOT NEEDS REVISION. I DON'T SEE ERROR BARS (HOW WAS STATISTICS DONE FOR THE RDFs?).I SUBMITTED NEW RDF CALCULATIONS CENTERED AROUND tRNA BECAUSE I DIDN'T FIND THE XVG FILES THAT WERE USED TO GENERATE THIS PLOT}}
\label{fig:tRNA_aggregation}
\end{figure}

@@

\caption{A) Solvent accessible surface area (sasa) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars). The average sasa for each protein in single molecule simulation is over number of chains of proteins and three replica. The average sasa for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica. The error bars show the standard errors. B) Root mean squared displacement (rmsd) for 8 proteins from single molecule simulation (blue bars) and cytoplasm simulation (orange bars). The average rmsd for each protein in single molecule simulation is over number of chains of proteins and three replica. The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica. The error bars show the standard errors.\textbf{\color{red}{THE MEASUREMENT NAME, SASA AND RMSD, IS MISSING IN THE Y-AXIS OF BOTH PLOTS}}}
\label{fig:structural_integrity_chain}
\end{figure}

@@

\begin{figure}[H]
\includegraphics[scale=0.5]{rmsf_2cols.pdf}
\caption{Root mean squared fluctuations (rmsf) for 8 crowders from single molecule simulation (blue bars) and cytoplasm simulation (orange bars). x-axis shows the residues for each chain of proteins. The average rmsf for each protein in single molecule simulation is over number of chains of proteins and three replica. The average rmsd for each protein in cytoplasm simulation are over number of chains of proteins, number of appearance and three replica. The error bars show the standard errors.\textbf{\color{red}{CAN YOU ADD THE PROTEIN NAME AS A LABEL INSIDE EACH PLOT? I THINK THE VARIABLE NAME SHOULD BE IN THE Y-AXIS (RMSF NM2)}}}
\label{fig:rmsf}
\end{figure}

@@

It has been demonstrated that each of the force fields that are most widely used to perform molecular dynamics simulations of macromolecules have different results for highly concentrated amino acid solutions ~ \cite{andrews2013molecular}. Previous all-atom simulations of high scale cytoplasm models used parameters from the CHARMM family of force fields ~\cite{yu2016biomolecular,mcguffee2010diffusion}. In this work we use, for the first time, the AMBER force fields to study crowding effects on a cytoplasm model. All-atom explicit solvent molecular dynamics simulations have been reported for large-scale cytoplasm models of \textit{Mycoplasma genitalium} composed by 103 and 12 million atoms that were simulated for 20 and 140 \SI{}{\nano\second}, respectively ~\cite{yu2016biomolecular}. The cytoplasm model described in this work has 1.5 million atoms and was simulated for the total of \SI{3}{\micro\second}, an unprecedented time scale to study crowding effects on a heterogeneous cytoplasm model in atomic detail.
